---
title: "Complications - Hospital"
author: "Shu-Han Chang"
date: "10/9/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```
### Introduction

I will analyze the Complications-Hospital dataset, which is an official dataset used on the Medicare.gov Hospital Compare Website provided by the Centers for Medicare & Medicaid Services. This dataset allows you to compare the quality of care on complications at over 4,000 Medicare-certified hospitals across the country. It is from https://data.medicare.gov/data/hospital-compare. It was accessed on October 9, 2017 by clicking on the "Export - Download As CSV for Excel" button. I used the library dplyr in the analysis.

### Data Acquisition and Selection

After downloading the csv file, I imported the dataset into R and kept only the columns needed for analysis. The file includes 81804 rows and 19 variables, which includes:

1. Provider ID
2. Hospital Name
3. Address
4. City
5. State
6. ZIP Code
7. County Name
8. Phone Number
9. Measure Name
10. Measure ID
11. Compared to National
12. Denominator
13. Score
14. Lower Estimate
15. Higher Estimate
16. Footnote
17. Meausure Start Date
18. Measure End State
19. Location

```{r getdata} 
library(readr)
Complications_and_Deaths_Hospital <- read_csv("C:/Users/linds/OneDrive/Fall 2017/Fall 2017/STAT-612 R/HW/HW4/Complications_and_Deaths_-_Hospital.csv", na = "Not Available")

#calculate the size of the dataset
nrow(Complications_and_Deaths_Hospital)
ncol(Complications_and_Deaths_Hospital)

#Create a new dataset, rename the variables and keep the variables needed for analysis
comps <- Complications_and_Deaths_Hospital %>%
          rename(hospital = 'Hospital Name',
                 compare = 'Compared to National',
                 lower = 'Lower Estimate',
                 higher = 'Higher Estimate',
                 measure_id = 'Measure ID',
                 state = State,
                 score = Score,
                 city = City,
                 measure_name = 'Measure Name') %>%
          filter(!is.na(score)) %>% #choose those that don't have na
          select(hospital, compare, lower, higher, measure_id, state, city, score, measure_name)
```
After filtering the records with no score and selecting the variables needed for analysis, there are 48281 records and 9 variables left in the new dataset "comp".

For the complication "Broken hip from a fall after surgery", I selected the data and write out a csv file, which includes hospital name, city, state, compared to national, score, lower estimate, and higher estimate.

After filtering the dataset with the measure ID of "Broken hip from a fall after surgery", the resulting file has 2850 records for each of the 7 variables that I chose.
```{r}
comps_hip <- comps %>% 
  filter(measure_id == "PSI_8_POST_HIP") %>% 
  select(hospital, city, state, compare, score, lower, higher)

write.csv(comps_hip, file = "comps_hip.csv",row.names = FALSE, na="")
```

### Data Analysis

#### Find the best hospital in state for a specific complication

To figure out if a high score or low score is better, I first found the national rate for the complication we selected, looked at the hospitals with rates below the national average, and arranged it by score or higher estimate.
```{r}
comps_hip %>%
  select(score,compare,lower,higher) %>%
  arrange(score)

comps_hip %>%
  select(score,compare,lower,higher) %>%
  arrange(higher)
```
According to the results above, the best hospitals should be judged as those with the lowest high estimate score.

Based on the results above, I wrote a best function that will take the id of a complication and a state as inputs, and return the hospital in the specified state with the lowest high estimate score for the complication.

```{r error=TRUE}
best <- function(ST, Complication) {

  FindST <- comps$state == ST
    if (sum(FindST)==0) stop ("invalid state entered")

  FindComplication <- comps$measure_id == Complication
    if (sum(FindComplication)==0) stop ("invalid complication entered")

  FindBoth <- FindComplication & FindST
    if (sum(FindBoth)==0) stop ("No entries found for the requested complication/state combination")
   
  BestDataOut <-  comps %>%
    filter(state == ST, measure_id == Complication) %>%
    select(hospital, state, higher, score, measure_id) %>%
    arrange(higher, score) 
  
  BestDataOut <- BestDataOut[1,]  
  
}
```

To test the best function for both valid and invalid inputs, I inserted a few inputs into the function.
```{r error=TRUE}
(bestdata1 <- best("DC" , "PSI_4_SURG_COMP"))
(bestdata2 <- best("AP" , "PSI_4_SURG_COMP"))
(bestdata3 <- best("DC", "pneumonia"))
```

#### Find the best hospitals in state for a specific complication

I then ranked the top n hospitals in a state with a given complication, if the number n is not inserted in the function or if n is larger than the number of hospitals in the state, the function will rank all the hospitals in the state.
```{r error=TRUE}
rankhospitalsinstate <- function(ST, Complication, n) {

  FindST <- comps$state == ST
    if (sum(FindST) == 0) stop ("invalid state entered")

  FindComplication <- comps$measure_id == Complication
    if (sum(FindComplication) == 0) stop ("invalid complication entered")

  FindBoth <- FindComplication & FindST
    if (sum(FindBoth) == 0) stop ("No entries found for the requested state and complication combination") 
  StateData <- comps %>%
    filter(state == ST, measure_id == Complication) %>%
    select(hospital, state, higher, score) 

    if (missing(n)) {
      n <- length(StateData$state)
    }
  
    if (n > length(StateData$state)) {
      n <- length(StateData$state) 
    }

  StateData <- arrange(StateData, higher, score)
  knitr::kable(StateData[1:n,]) #:: <- specify which package you want to use with the function
                                #kable - create a table 
}
```

To test the function, I inserted "VA", "PSI_15_ACC_LAC",10 to check for the top 10 hospitals in Virginia with the complication ""Accidental cuts and tears from medical treatment". In addition to top 10 hospitals, I also tried inserting 500 hospitals and omitting the number of hospitals in the function.

```{r error=TRUE}
rankhospitalsinstate("VA","PSI_15_ACC_LAC",10)
rankhospitalsinstate("VA","PSI_15_ACC_LAC",500)
rankhospitalsinstate("VA","PSI_15_ACC_LAC")
```

#### Rank all hospitals in the country for a specific complication

Lastly, I ranked the top hospitals in the country for a complication by selecting the records for all hospitals with a specific complication.

```{r error=TRUE}
rankallhospitals <- function(Complication, num=1) {
  #if don't provide the value of "num" in the function to overwrite it, 1 is the default
  RankHospital <- comps %>%
    filter(measure_id == Complication) %>%
    select(hospital, state, higher, score, measure_id) %>%
    arrange(higher,score)
  
  RankHospital[1:num,]
}
```

To test the function, I checked the top 10 hospitals in the country with the complication "Serious blood clots after surgery". In addition to that, I also tested the function by omitting the number of hospitals in the function, with the complication "A wound that splits open after surgery on the abdomen or pelvis".

```{r error=TRUE}
rankallhospitals("PSI_12_POSTOP_PULMEMB_DVT",10)
rankallhospitals("PSI_14_POSTOP_DEHIS")
```

### System Information

```{r}
sessionInfo()
```